Kindly Bent to Free Us

### 2020-05-28 kindly/motivation.tex
low level stuff added

If we want to work on borrows of linear and affine resources, then the
typing gets more involved because
we must make sure those borrows are not leaked in the result.
We obtain the following signature for @bfold@, the borrowing fold operation:
\begin{lstlisting}[numbers=none]
val bfold : ('b:'k),('k <= aff_r) => (&(aff_r1,'a) -> 'b -{aff_r1}> 'b) -> &('k_1,'a Array.t) -> 'b -{'k_1}> 'b
\end{lstlisting}

The folded function receives a shared borrow of the element in the array.
The typing of the callback ensures
that this borrow is neither captured nor returned by the function.
%
This encapsulation is implemented with a universally quantified \emph{kind index variable} $r$.
The signature prescribes the type @&(aff_r1,'a)@ for the
shared borrow of the resource with an affine kind at region nesting $r+1$. The return
type of the callback is constrained to kind @'k <= aff_r@
so that the callback certainly cannot return the borrowed argument.
%
The input of the fold is a shared borrow of the array,
which ensures that we have the right to share borrows of the inner content and
make multiple concurrent folds.

@@

>>>>

If we want to work on borrows of linear and affine resources, then the
typing gets more involved because
we must make sure those borrows are not leaked in the result.
We obtain the following signature for @bfold@, the borrowing fold operation:
\begin{lstlisting}[numbers=none]
val bfold : ('b:'k),('k <= lin_r) => (&(aff_r1,'a) -> 'b -{aff_r1}> 'b) -> &('k_1,'a Array.t) -> 'b -{'k_1}> 'b
\end{lstlisting}

The folded function receives a shared borrow of the element in the array.
The typing of the callback ensures
that this borrow is neither captured nor returned by the function.
%
This encapsulation is implemented with a universally quantified \emph{kind index variable} $r$.
The signature prescribes the type @&(aff_r1,'a)@ for the
shared borrow of the resource with an affine kind at region nesting $r+1$. The return
type of the callback is constrained to kind @'k <= lin_r@. The
important part of this constraint is the $_r$ index, which ensures
that the callback cannot return the borrowed argument from the more
deeply nested scope. 
%
The input of the fold is a shared borrow of the array,
which ensures that we have the right to share borrows of the inner content and
make multiple concurrent folds.

@@

To close this discussion, let's see which primitives are needed to implement
functions like @bfold@ and @iter_mut@. The obvious sequential implementation of
both functions boils down to a loop over the index range of the array:
\begin{lstlisting}
let rec bfold_helper f a z i =
  if i >= 0
  then bfold_helper f &a (f (get_sb (&a, i))) (i - 1)
  else z
let bfold f a z =
  bfold_helper f a z (length &a - 1)
\end{lstlisting}
\begin{lstlisting}
let rec iter_helper f a i =
  if i >= 0
  then (f (get_eb (&!a, i)); iter_helper f &!a (i - 1))
  else ()
let iter_mut f a =
  iter_mut f a (length &a - 1)
\end{lstlisting}
Observe that we are forced to use different primitives
\begin{itemize}
\item @get_sb : &('r, 'a Array.t) -> int -> &('r, 'a)@ to get a
  shared borrow from a shared borrow of an array and
\item @get_eb : &!('r, 'a Array.t) -> int -> &('r, 'a)@ to get an
  exclusive borrow from an exclusive borrow of  an array.
\end{itemize}
These primitives have the same underlying implementing (the same as
@get@ from \cref{sec:imper-progr}). Their types arise from the
intuition that the borrow of a structure should entitle to borrows of
its substructures, roughly, the borrow of an array could be considered as
array of borrows.

%%% PJT: TBC
